Algebraic File Synchronization: Adequacy and Completeness

### 2017-11-16 p2/axioms.tex
Minor improvements
TeX improvements
Command categories

\section{Investigating Command Pairs}\mlabel{section_axioms}

@@

\begin{mydef}[Construction pair]
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{gather*}
            \cbdaa{\nn}\cc  \cbfaa{n} \\
            \cbdaa{\nn}\cc  \cbdaa{n} \\
            \cfdaa{\nn}\cc  \caaa{\empt}{\valff}{n} \\
            \cfdaa{\nn}\cc  \cbdaa{n}
   \end{gather*}
\end{mydef}

@@

\begin{mydef}[Destruction pair]
A pair of commands on nodes $n$ and $\nn$ is a destruction pair if $\parent(n)=\nn$ and the commands are one
of the following:
   \begin{gather*}
            \cfba{n}\cc  \cdba{\nn} \\
            \cfba{n}\cc  \caaa{\vald}{\valff}{\nn} \\
            \cdba{n}\cc  \cdba{\nn} \\
            \cdba{n}\cc  \cdfaa{\nn}
   \end{gather*}
\end{mydef}

@@

\begin{mylem}\mlabel{incomparable_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are incomparable. In other words,
\[ \cxynv\indep\czwmv \Longleftrightarrow n\unrel m \]
if $\cxy\neq\czw$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
This is easy to show based on the output of the program that investigates the behavior of command pairs.
However, this proposition can also be derived from the \namecrefs{ax_separate_commute} already listed
in the following way.

>>>>

\section{Command Pairs and Sequences}\mlabel{section_axioms}

@@

\begin{mydef}[Construction pair]
A pair of commands on nodes $\nn$ and $n$ is a construction pair if $\nn=\parent(n)$ 
and the commands are one
of the following:
   \begin{align*}
            \cbda{\nn}&\cc  \cbfa{n} \\
            \cbda{\nn}&\cc  \cbda{n} \\
            \cfda{\nn}&\cc  \caaa{\empt}{\valff}{n} \\
            \cfda{\nn}&\cc  \cbda{n}
   \end{align*}
\end{mydef}

@@

\begin{mydef}[Destruction pair]
A pair of commands on nodes $n$ and $\nn$ is a destruction pair if $\parent(n)=\nn$ and the commands are one
of the following:
   \begin{align*}
            \cfba{n}&\cc  \cdba{\nn} \\
            \cfba{n}&\cc  \caaa{\vald}{\valff}{\nn} \\
            \cdba{n}&\cc  \cdba{\nn} \\
            \cdba{n}&\cc  \cdfa{\nn}
   \end{align*}
\end{mydef}

@@

\begin{mylem}\mlabel{incomparable_is_independent}
Two different commands that are not assertion commands are independent iff the nodes they change are incomparable. In other words,
\[ \cxynv\indep\czwmv \Longleftrightarrow n\unrel m \]
if $\cxynv\neq\czwmv$ and none of $\cxy$ and $\cxw$ is $\cbb$ or $\cdd$.
\end{mylem}
\begin{proof}
This is easy to show based on the output of the program that investigates the behavior of command pairs.
However, this proposition can also be derived from the \namecrefs{ax_separate_commute} already listed
in the following way.

### 2017-11-16 p2/comparisons.tex
Command categories

The former observation is supported by Berzan and Ramsey, who in \cite{CBNR} 
describe different general reconciliation policies.
The liberal (maximal) policy propagates all updates to all replicas where
the update does not break the filesystem, while a conservative policy
refrains from updating any node that is below a node with conflicting commands.
They show that the reconciliation algorithm in \cite{NREC} implements
an intermediate policy as one can easily construct two- or three-replica scenarios
where an update could clearly be propagated, but it is excluded.
To rephrase the example in \cite{CBNR}, consider the following two
update sequences that have been applied to replicas $\FS_A$ and $\FS_B$:
\begin{align*}
A&=\cbdaa{\nn}{\valdi}\cc\cbfaa{n}{\valf} \\
B&=\cbdaa{\nn}{\valdii}
\end{align*}
where $\parent(n)=\nn$ (\cite{NREC} did not require all directories to have the same value).
Clearly $\cbfaa{n}{\valf}$ could be applied to $\FS_B$, but it is not as
it must be preceded by creating the directory, which conflicts with the same command in $B$
due to the different value.
(Introducing a third replica which has not changed at all further complicates the picture.)
The current algorithm no longer needs to specify that there can be no conflicts
on preceding commands, and, to use the above terminology, implements a fully liberal policy.

>>>>

The former observation is supported by Berzan and Ramsey, who in \cite{CBNR} 
describe different general reconciliation policies.
The liberal (maximal) policy propagates all updates to all replicas where
the update does not break the filesystem, while a conservative policy
refrains from updating any node that is below a node with conflicting commands.
They show that the reconciliation algorithm in \cite{NREC} implements
an intermediate policy as one can easily construct two- or three-replica scenarios
where an update could clearly be propagated, but it is excluded.
To rephrase the example in \cite{CBNR}, consider the following two
update sequences that have been applied to replicas $\FS_A$ and $\FS_B$:
\begin{align*}
A&=\caaa{\empt}{\valdi}{\nn}\cc\caaa{\empt}{\valf}{n} \\
B&=\caaa{\empt}{\valdii}{\nn}
\end{align*}
where $\parent(n)=\nn$ (\cite{NREC} did not require all directories to have the same value).
Clearly $\caaa{\empt}{\valf}{n}$ could be applied to $\FS_B$, but it is not as
it must be preceded by creating the directory, which conflicts with the same command in $B$
due to the different value.
(Introducing a third replica which has not changed at all further complicates the picture.)
The current algorithm no longer needs to specify that there can be no conflicts
on preceding commands, and, to use the above terminology, implements a fully liberal policy.

### 2017-11-16 p2/def-commands.tex
Command categories

\begin{mydef}[Filesystem commands]
Filesystem commands are partial endofunctions on filesystems:
\[ \alpha: \setfs \nrightarrow \setfs \]
A command $\alpha$ is not defined on filesystems where it returns an error.
The commands are represented using triplets of the form
\[ \alpha = \caaaa{x}{y}{n}, \]
where $x,y\in\setv$ and $n\in\setn$.
The equivalence class $[x]$ is the \emph{input type} of the function;
$y$ is its \emph{output value} implicitly specifying the \emph{output type} $[y]$,
and $n$ is the node the command is applied to.
\end{mydef}
For example, $\cbfaa{n}$ represents \texit{create}$(n,\valf)$,
and $\cdba{n}$ represents \textit{rmdir}$(n)$.

@@

We note that a command or a sequence of commands is applied to a filesystem
by prefixing the command or sequence to it, for example: $\cbrk\aFS$, $\cbdaa{n}\aFS$, 
or $S\aFS$ if $S$ is a sequence of commands.

@@

\begin{mydef}[Effect of commands]
The effect of the command $\cxynv$ is defined as follows:
\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\textrm{if $\FS(n)]\neq[x]$,}\\
   \fsbroken &\textrm{if $\fsreplacement{\FS}{n}{y}$ violates the tree property,}\\
   \fsreplacement{\FS}{n}{y} &\textrm{otherwise.}
   \end{cases}
\end{mydef}
In other words, the command $\cxynv$ breaks a filesystem if the input type $[x]$
does not match the type of the value in the filesystem at $n$, or if after replacing
the value at $n$ with $y$, the resulting filesystem ceases to satisfy the tree property.

>>>>

\begin{mydef}[Filesystem commands]
Filesystem commands are partial endofunctions on filesystems:
\[ \alpha: \setfs \nrightarrow \setfs \]
A command $\alpha$ is not defined on filesystems where it returns an error.
The commands are represented using triplets of the form
\[ \alpha = \caaaa{x}{y}{n}, \]
where $x,y\in\setv$ and $n\in\setn$.
The equivalence class $[x]$ is the \emph{input type} of the function;
$y$ is its \emph{output value} implicitly specifying the \emph{output type} $[y]$,
and $n$ is the node the command is applied to.
\end{mydef}
For example, $\cbfa{n}$ represents \textit{create}$(n,\valf)$,
and $\cdba{n}$ represents \textit{rmdir}$(n)$.

@@

We note that a command or a sequence of commands is applied to a filesystem
by prefixing the command or sequence to it, for example: $\cbrk\aFS$, $\cbda{n}\aFS$, 
or $S\aFS$ if $S$ is a sequence of commands.

@@

\begin{mydef}[Effect of commands]
The effect of the command $\cxynv$ is defined as follows:
\[
\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\textrm{if $[\FS(n)]\neq[x]$,}\\
   \fsbroken &\textrm{if $\fsreplacement{\FS}{n}{y}$ violates the tree property,}\\
   \fsreplacement{\FS}{n}{y} &\textrm{otherwise.}
   \end{cases}
\]
\end{mydef}
In other words, the command $\cxynv$ breaks a filesystem if the input type $[x]$
does not match the type of the value in the filesystem at $n$, or if after replacing
the value at $n$ with $y$, the resulting filesystem ceases to satisfy the tree property.

\bigskip
\noindent
Depending on their input and output types, there are nine types of commands,
which we separate into four categories depending on the overall effect of the command.
Constructions commands extend the filesystem,
while destruction commands shrink it;
the replacement command replaces a file value with a (presumably different) file value,
and assertion commands either break the filesystem, or assert the type of the value at a node.
\begin{mydef}[Command categories]
Depending on their input and output types, filesystem commands
belong to exactly one of the following four categories:
\begin{itemize}
\item[]\emph{Construction commands:}
    $\cbfa{n}$, $\cbda{n}$ and $\cfda{n}$;
\item[]\emph{Destruction commands:}
    $\cdfa{n}$, $\cdba{n}$ and $\cfba{n}$;
\item[]\emph{Assertion commands:}
    $\cbba{n}$ and $\cdda{n}$;
\item[]\emph{Replacement command:}
    $\caaa{\valfx}{\valff}{n}$.
\qedhere
\end{itemize}
\end{mydef}
